\textsf{Are large language models superhuman chemists?}

### 2024-10-18 src/tex/appendix.tex

% \begin{figure}[htb]
%     \centering
%     \includegraphics{figures/confidence_score_distributions.pdf}
%     \caption{\textbf{Distribution of confidence scores reported by \glspl{llm}.} \Glspl{llm} show different distributions of confidence scores. The confidence scores are reported on an ordinal scale from 1 to 5, with 1 indicating low confidence and 5 indicating high confidence. The bar plots show how many questions were answered with each confidence score.}
%     \label{fig:confidence_score_distributions}
%     \script{plot_confidence_score_distributions.py}
% \end{figure}

>>>>

Our analysis of both log probabilities and prompting confidence reveals distinct calibration behaviors across different language models. 
GPT-4o demonstrates an overconfident tendency, often assigning high probabilities even to incorrect answers. 
However, when GPT-4o displays high confidence, it accurately predicts correct answers approximately 80\% of the time. In contrast, Llama-3.1-Instruct confidence distribution is more evenly spread, with a majority of predictions centered around 0.5 probability. High-confidence predictions from Llama are less frequent compared to GPT-4o, and unlike GPT-4o, high confidence does not necessarily correlate with a higher chance of correct answers.
\begin{figure}[htb] 
    \centering
    \includegraphics{figures/calibration_plot_overall_filtered.pdf}
    \caption{\textbf{Calibration between linear probability of \glspl{llm} and correctness.} For this analysis we obtained the linear probability from the logprobs of the models. Only logprobs of the tokens corresponding to the answers were considered. Linear proabability was computed by taking exponential of logprobs (for sequences with multiple tokens, values were multiplied).  The plot shows the average predicted probability and the fraction of correct answers for each bin of linear probabilities. The ideal scenario is a diagonal line, indicating perfect calibration where the model's confidence aligns perfectly with the actual correctness. The ECE (Expected Calibration Error) value quantifies the overall calibration performance, with a lower ECE indicating better calibration.}
    \label{fig:confidence_score_distributions}
    \script{plot_logprobs.py}
\end{figure}

### 2024-10-18 src/tex/ms.tex

\begin{figure}[!htb]
    \centering
    \includegraphics[width=\textwidth]{figures/cb_corpus_v2.pdf}
    \caption{\textbf{Distribution of topics and required skills.} This circular plot illustrates the distribution of questions across various chemistry topics, along with the primary skills required to address them. The topics were manually classified, showing a varied representation across different aspects of chemistry. Each topic is associated with a combination of three key skills: Calculation, Reasoning, and Knowledge, as indicated by the colored bars. \chembench samples diverse topics and diverse skills, setting a high bar for \glspl{llm} to demonstrate human-competitive performance across a wide range of chemistry tasks.}
    \label{fig:corpus}
\end{figure}

@@

% We also gain insight into the models' struggles with chemical reasoning tasks when we examine their performance as a function of molecular descriptors.
% If the model would answer questions after reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
% However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{tab:correlation_coefficients}).
% This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on retrieval of fragments of the training data.
% Those observations mirror recent findings that \glspl{llm} struggle to \enquote{reverse} facts they have seen in training (i.e., generalize from \enquote{A has feature B} to \enquote{B is a feature of A}).\autocite{berglund2023reversal, zhu2023physics, allen2023physics, golovneva2024reverse}

% It is important to note that the model performance for some topics, however, is underestimated in the current evaluation.
% This is because models provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
% For instance, models might refuse to provide answers about cyanides.
% To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
% This is facilitated with the tooling \chembench provides, thanks to which contributors can automatically add new models in an open science fashion.

>>>>

\begin{figure}[!htb]
    \centering
    \includegraphics[width=\textwidth]{figures/cb_corpus_v3.pdf}
    \caption{\textbf{Distribution of topics and required skills.} This circular plot illustrates the distribution of questions across various chemistry topics, along with the primary skills required to address them. The topics were manually classified, showing a varied representation across different aspects of chemistry. Each topic is associated with a combination of three key skills: Calculation, Reasoning, and Knowledge, as indicated by the colored bars. \chembench samples diverse topics and diverse skills, setting a high bar for \glspl{llm} to demonstrate human-competitive performance across a wide range of chemistry tasks.}
    \label{fig:corpus}
\end{figure}

@@

 We also gain insight into the models' struggles with chemical reasoning tasks when we examine their performance as a function of molecular descriptors.
 If the model would answer questions after reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
 However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{tab:correlation_coefficients}).
 This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on proximity of the molecules to the training data.\autocite{mccoy2023embersautoregressionunderstandinglarge}

It is important to note that the model performance for some topics, however, is slightly underestimated in the current evaluation.
 This is because models provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
 For instance, models might refuse to provide answers about cyanides. Statistics of the frequency of such refusals are shown in \Cref{tab:refusals}.
 To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
 This is facilitated with the tooling \chembench provides, thanks to which contributors can automatically add new models in an open science fashion.
